---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(SEERreadr)
```
# SEERreadr

<!-- badges: start -->
<!-- badges: end -->

A small package for reading SEER fixed width files. 

<!-- README start -->

## Installation

**SEERreadr** can be installed from GitHub with

```r
# install.packages("remotes")
remotes::install_github("gerkelab/SEERreadr", upgrade = FALSE)
```

## What does this package do?


### Read SEER Data Files

The main workhorse of this package is `seer_read_fwf()`. 
This function wraps `readr::read_fwf()` to import the SEER fixed-width ASCII data files, using the column names and field width definitions in the 
[SEER SAS script](https://seer.cancer.gov/manuals/read.seer.research.nov2017.sas). 

The data files are available from the 
[SEER Data & Software page](https://seer.cancer.gov/data-software/), 
where users must request access prior to downloading. 
The SAS script is included in the file download, or avilable online. 
The online version is used by `seer_read_fwf()`, but a local version can be specified in the helper function `seer_read_col_positions("local_file.sas")`.

```r
library(SEERreadr)
x <- seer_read_fwf("incidence/yr1973_2015.seer9/MALEGEN.TXT")
```

### Recode SEER Variables

Two additional functions are provided to help recode the SEER data. 
`seer_recode()` uses the  `seer_data_dictionary` data provided in this package to automatically recode all variables with a one-to-one correspondence, for example:

```{r}
seer_data_dictionary$SEX
```

The package also includes the function `seer_rename_site_specific()` that can be used to replace the site-specific variables with their corresponding labels, formatted appropriately to serve as variable names.
As an example, CSSSF variables for Breast cancer would be renamed according to the following table.

```{r echo=FALSE}
seer_data_dictionary$CSSSF %>% 
  dplyr::filter(`Schema Name` == "Breast") %>% 
  dplyr::mutate(label = snakecase::to_snake_case(label)) %>% 
  dplyr::select(`Original Variable` = variable, `New Variable Name` = label) %>% 
  knitr::kable()
```

## Thanks

Thank you to [Vincent Major](https://github.com/vincentmajor) for making available the scripts in [SEER_read_fwf](https://github.com/vincentmajor/SEER_read_fwf), which provided a foundation for this package.